### 2013-12-06 paper/main.tex
finished documentation of Direct Design method
The Direct Design method has been documented which
includes the steps needed to calculate D(z) and K.

\usepackage{tikz}
\usetikzlibrary{calc,arrows,positioning,shadows}
%
% control system tikz objects
\tikzset{shadow/.style={drop shadow,fill=white}}
\tikzset{sum/.style={circle,draw,very thick,shadow,minimum size=6mm}}
\tikzset{sumnp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumpp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$+$}} }
\tikzset{sumpn/.style={sum,
				label={{below left,xshift=0.0mm,yshift=1.0mm}:$+$},
				label={{above right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumxpxp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-0.0mm,yshift=-1.0mm}:$+$}} }
\tikzset{gain/.style={rectangle,draw,very thick,inner sep=2mm,shadow}}

@@

% {{{ fig:dd_plot2
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/dd_plot2}
\end{center}
\caption{Output response of controller built using Direct Design
compared to no control.  For no control this is the same plot
as in Figure \ref{fig:em_plot4} which stabilizes after a second.}
\label{fig:dd_plot2}
\end{figure}
% }}}

@@

% {{{ Engine Model Matlab Source
\clearpage
\section{Engine Model Matlab Source}
\label{app:em_matlab}
\label{app:rpmtime_plot_source}

@@

@@

\lstinputlisting[
	caption={Matlab source used to plot the controller built using
	Direct Design compared to no control.}
	]{../octave/dd_plot2.m}

>>>>

\usepackage{tikz}
\usetikzlibrary{calc,arrows,positioning,shadows}
%
% control system tikz objects
\tikzset{shadow/.style={drop shadow,fill=white}}
\tikzset{sum/.style={circle,draw,very thick,shadow,minimum size=5mm}}
\tikzset{sumnp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumpp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$+$}} }
\tikzset{sumpn/.style={sum,
				label={{below left,xshift=0.0mm,yshift=1.0mm}:$+$},
				label={{above right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumxpxp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-0.0mm,yshift=-1.0mm}:$+$}} }
\tikzset{gain/.style={rectangle,draw,very thick,inner sep=2mm,shadow}}

@@

There are several characteristics of this Direct Design method.
The controller is built entirely in the discrete domain.
If the plant ($G(z)$) is continuous it is converted to discrete
using a Zero Order Hold.
In this case it is already discrete.
It is a pole placement technique where the desired roots of the
entire system are specified by the designer.
The controller values for $\beta$ and $\alpha$ are found using the
Diophantine equation along with a Sylvester matrix.
And the gain ($K$) is chosen such that the desired limit is reached.
Figure \ref{fig:ddc} shows the structure of the system.

% {{{ fig:ddc
\begin{figure}[h!]

	\draw [<-] (K) -- ++(-20mm,0) node[above]{$u$};
	\draw [->] (K) -- (sum1);
	\draw [->] (sum1) -- (Gz);
	\draw [<-] (sum1) |- (Dz);
	\draw [->] (Gz) -- ++(35mm,0) node[above]{$y$};
	\draw [->] (Gz) -- ++(20mm,0) |- (Dz.east);

	% add B/A labels
	\node [right of=Gz,node distance=0cm,xshift=9mm,yshift=6mm] {$\dfrac{B}{A}$};
	\node [right of=Dz,node distance=0cm,xshift=9mm,yshift=6mm] {$\dfrac{\beta}{\alpha}$};
\end{tikzpicture}
\end{center}

\caption{Direct Design system where $K$ is the scaling input, $G$ is
the plant and $D$ is the controller.}
\label{fig:ddc}
\end{figure}
% }}}

In this case the roots were not given and no deterministic way
was available to find them.
So trial and error was used.

\begin{align*}
	\text{roots} &: z = 0.6 \pm 0.4i
\end{align*}

Since the plant is already in discrete form it is not necessary
to use a Zero Order Hold.
The engine model, repeated from Equation \ref{eq:lems}, is shown below.
\begin{align*}
G(z) &= \frac{43.38}{z^3 - 1.838 z^2 + 0.8442 z + 0.02375} \\
	&= \frac{B(z)}{A(z)} = \frac{b_0 z^3 + b_1 z + b2}
								{a_0 z^3 + a_1 z + a2}
\end{align*}

Next it must be arranged to place it in the form of the Diophantine Equation.

From the block diagram the equation describing the system is found.
\begin{align*}
	\frac{y}{h} &= H \\
	H &= K \frac{G}{1 + D G} \\
	H &= K \frac{B \alpha}{A \alpha + B \beta}
\end{align*}

From the roots $D$ is constructed.
\begin{align*}
	D(z) &= (0.6 + 0.4i)(0.6 - 0.4i) \\
	D(z) &= z^2 - 1.2z + 0.520
\end{align*}

Recall the form of the Diophantine Equation where the order is $n$.
\[
\begin{array}{ccccccc}
	\alpha(z) &A(z) &+ &\beta(z) &B(z) &= &D \\
	  (n-1)   &(n)  &  & (n-1)  &(n) & &(2n-1)
\end{array}
\]

Since $A(z)$ is a third order polynomial $n = 3$.
This requires that $D(z)$ be a fifth order polynomial ($2n-1$).
The order is increased by adding zeros.

\begin{align*}
	D(z) &= z^5 - 1.2z^4 + 0.520z^3 \\
		 &= d_0z^5 + d_1z^4 + d_2z^3 + d_3z^2 + d_4z + d_5
\end{align*}

This also requires that $\alpha$ and $\beta$ are
second order polynomials ($n-1$).

\begin{align*}
	\frac{\beta}{\alpha} &= \frac{\beta_0 z^2 + \beta_1 z + \beta_2}
								{\alpha_0 z^2 + \alpha_1 z + \alpha_2}
\end{align*}

To solve the Diophantine Equation

\begin{align*}
	M &= E^{-1} D \\
	\left[
	\begin{array}{l}
		\alpha \\
		\beta
	\end{array}
	\right] &= E^{-1} D \\
	\left[
	\begin{array}{l}
		\alpha \\
		\beta
	\end{array}
	\right] &= 
	\left[
	\begin{array}{l}
		\alpha_2 \\
		\alpha_1 \\
		\alpha_0 \\
		\beta_2 \\
		\beta_1 \\
		\beta_0
	\end{array}
	\right]
\end{align*}

a third order Sylvester Matrix is required.

\[
E=
\begin{bmatrix}
	a_3 & 0 & 0 & b_3 & 0 & 0 \\
	a_2 & a_3 & 0 & b_2 & b_3 & 0 \\
	a_1 & a_2 & a_3 & b_1 & b_2 & b_3 \\
	a_0 & a_1 & a_2 & b_0 & b_1 & b_2 \\
	0 & a_0 & a_1 & 0 & b_0 & b_1 \\
	0 & 0 & a_0 & 0 & 0 & b_0 \\
\end{bmatrix}
\]

Combining these parts and solving for $\alpha$ and $\beta$ results
in a solution for $D(z)$.

\begin{align*}
	\Aboxed{
	\frac{\beta}{\alpha} &= D(z) = \frac{0.02297 z^2 - 0.01686 z - 0.0004643}
								{z^2 + 0.6379 z + 0.8482}
	}
\end{align*}

\begin{samepage}
The gain ($K$) still needs to be found and this can be accomplished
using the limit.
\begin{align*}
\lim_{z \to 1} &K \frac{B \alpha}{A \alpha + B \beta} = 1\\
\lim_{z \to 1} &K \frac{(43.38) (z^2 + 0.6379 z + 0.8482)}
	{(z^5 - 1.2z^4 + 0.520z^3)} = 1 \\
K &= \frac{(1 - 1.2 + 0.520)}{(43.38)(1 + 0.6379 + 0.8482)} \\
\Aboxed{ K &= 0.0029671 }
\end{align*}
\end{samepage}

This solution results in the response shown in Figure \ref{fig:dd_plot2}.
It is a dramatic improvement compared to no control.
The rise time, percent overshoot, and settling time are all improved.
This result is particularly interesting because the chosen roots were a guess.
It is likely that further improvement could be made by optimizing the roots.

% {{{ fig:dd_plot2
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.6]{../octave/dd_plot2}
\end{center}
\caption{Output response of controller built using Direct Design
compared to no control.  For no control this is the same plot
as in Figure \ref{fig:em_plot4} which stabilizes after a second.
The Matlab source performing these calculations and producing this
plot is given in Appendix \ref{app:dd_matlab}.}
\label{fig:dd_plot2}
\end{figure}
% }}}

While these results were positive it is not obvious whether it is withing
the mechanical limits of a real engine.
In particular the control input should be limited to a range from zero to one.
Further investigation is needed to determine whether these limits
are being exceeded.
There is no doubt, however, that the addition of a controller
improved performance.
If the limits are being exceeded a different set of roots would have to
be chosen which may be less optimal.

@@

% {{{ Engine Model Matlab Source
\clearpage
\section{Engine Model Matlab Source}
\label{app:em_matlab}

@@

% {{{ General Matlab Functions
\clearpage
\section{General Matlab Functions}
\label{app:gf_matlab}

This sections includes Matlab source for functions which are
used among several methods.

\lstinputlisting[
	caption={Matlab source of function to calculate a Sylvester Matrix.} 
	]{../octave/sylvester.m}

\lstinputlisting[
	caption={Matlab source of function to calculate zero time.} 
	]{../octave/zerotime.m}

\clearpage
\lstinputlisting[
	caption={Matlab source of modified lsim() function.} 
	]{../octave/lsim1.m}

@@

Matlab source code used to build a controller using Direct Design
and plot the output.
The \verb+engine_model+ is given in Appendix \ref{app:em_matlab}
and other functions are given in Appendix \ref{app:gf_matlab}.

\lstinputlisting[
	caption={Matlab source of function is to build a controller using
		Direct Design.}
	]{../octave/direct_design.m}

\clearpage
\lstinputlisting[
	caption={Matlab source used to plot the controller built using
	Direct Design compared to no control.}
	]{../octave/dd_plot2.m}
